Analysis \& Visualization of Call Center Data using Python

### 2018-08-27 paper.tex
worked on analysis section

@@

@@

	\subsection{Discuss packages and analysis plan of attack???}

@@

@@

@@

@@

\begin{thebibliography}{6}

\bibitem{}

\end{thebibliography}

\includepdf[pages=-]{311CallCenterDashboard}

>>>>

The goal of this analysis and visualization task is to, at its core, provide the city of South Bend useful and actionable knowledge about its call center.  The city was most curious about trends based on time, 
%I dont like that word
departments, and topics.  Therefore, this analysis has four major analytical focuses:  month of year, day of the week, department, and topic.  
%I hate everything here
In order to illustrate the finding in the most information-rich way possible, an interactive graphical dashboard was chosen as the primary means of communication.  This dashboard is able to be sent throughout the city management conveniently, as it is simply a self-contained HTML file.  Neither Python nor the data sources are necessary to view and interact with this visualization.

@@

	\subsection{Programming \& Packages}

This entire analysis was conducted in Anaconda 5.2 running Python 3.6.4.  The Python packages used throughout this analysis are all contained within the Anaconda distribution.  These include Numpy, Pandas, Matplotlib, Scipy, Seaborn, and Bokeh.\cite{python}\cite{anaconda}  The programming and analysis was all conducted using Windows 10 OS.  The methods and output, however, should be able to work on Python 3.6 (with the required packages) on any OS.  The code, however, is not backwards-compatible to Python 2.x.

	\subsection{Data}

@@

@@

An analysis of call volume by month was conducted using the daily data, which provides aggregate totals by day for a 2-year period. The relevant data was manipulated by taking the mean of the daily call volume for every record in each month.  Upon completion, the data was then plotted in a bar chart format, allowing the user to quickly identify any differences.  This chart can be viewed in Figure 1.

\begin{figure}[h]
  \includegraphics[scale=.27]{month_bar.png}
  \caption{Plot of average daily call volume by month}
\end{figure}

As an exercise in thoroughness, these findings were subjected to a two-sided t-test for statistical significance.  The daily call volumes for each month were isolated and stored as lists within a Python dictionary.  The scipy.stats package was used to conduct the test.  For the purposes of this test, the null hypothesis stated that ``the two months have identical expected values."  For the null hypothesis to be disproved, a standard threshold p-value of 0.05 was used.  The heatmap in Figure 2 shows the p-value between each pairing of months.  From this information, the city can see which months are significantly busier/quieter than others.  By subjecting the data to such a test, important points can be quickly identified that may or may not be obvious in Figure 1.

\begin{figure}[h]
	\includegraphics[scale=.65]{Heatmap.png}
	\caption{P-values resulting from t-test on daily call volume by month}
\end{figure}

@@

One of the primary tasks posed by the city was to explore call trends on a departmental level.  This was conducted primarily using the topic data, as the data contained accurate and complete information regarding departments.  In addition to being the most recent, this data also utilizes the new department/topic layout that the city has transitioned to.  This format includes the knowledge base articles described earlier.  In this native data format, the department titles were often followed by the characters ``- KB Team".  Since leaving this suffix appended to each title would only visually detract from the dashboard, the string was removed.  This process can be seen below:\\

  \begin{Verbatim}[fontsize=\small]
  for i in range(len(dept)):
    n = dept[i]
    dept[i] = n.replace(" - KB Team", "")
  \end{Verbatim}

Additionally, the time format in the data file was read in as a string.  Upon investigating, the data was found to have an inconsistent format.  For example, duration for one record was in the form ``2:34" for two minutes, thirty-four seconds.  However, for longer calls it could be in the form ``1:15:56" for one hour, fifteen minutes, and fifty-six seconds.  While beneficial visually, this would not function for any analysis.  This data was converted into seconds, stored as a numeric data type, using the following function:

  \begin{Verbatim}[fontsize=\small]

def string_to_sec(df, col_name):
    ''' 
    This function accepts a dataframe and 
    column name (in quotes)
    It converts the column from XX:XX:XX to
     seconds and returns a list.
    Pandas must be imported and 
    aliased as pd
    '''
    new = []
    for i in range(len(df[col_name])):
        x = df.loc[i][col_name]
        if x.count(":") == 1:
            ''' For Min:sec '''
            y, z = x.split(":")
            y = pd.to_numeric(y)
            z = pd.to_numeric(z)
            q = (y * 60) + z
        elif x.count(":") == 2:
       '''For H/Min:Min/sec:Sec/Nothing '''
            h, y, z = x.split(":")
            if len(h) == 2:
                ''' This assumes if the
               first number is 2 digits,
               it is minutes'''
                h = pd.to_numeric(h)
                y = pd.to_numeric(y)
                q = (h * 60) + y
             elif len(h) == 1:
              '''Assuming first digit is 
                single, means hours'''
                h = pd.to_numeric(h)
                y = pd.to_numeric(y)
                z = pd.to_numeric(z)
                q = (h*3600) + (y*60) + z
             else:
                 q = np.NaN
         else:
              q = np.NaN
         new.append(q)
  return new
  \end{Verbatim}

This function returns a list, which is then evaluated for reasonableness and appended to the dataframe.  By doing this, every call record now has a new attribute for duration in seconds.

Calls by department were then analyzed and visualized by average duration and total volume.  As expected, the departments' respective ordering in each analysis differed.  As a more direct approach to the measure of total call volume, the total departmental call times were calculated.  Of all the approaches, this calculation of total call time may be the most helpful to the city, as this can directly show where time and budget is being spent.  This visualization can be seen in the pie chart in Figure 3.

\begin{figure}[h]
  \includegraphics[scale=0.3]{piechart.png}
  \caption{A visual of total volume by department}
 \end{figure}

@@

Analyzing calls by topic proved to be one of the most challenging and interesting aspects of this analysis.  The topic data was used again here due to its recentness and the homogeneity of topic format.  Among this dataset, 426 unique call topics existed.  The data was then processed to show total call count and average call duration by topics.  This data was then merged with a separate dataframe, indexed by topic with department as the only attribute.  Once completed, the resulting dataframe contained topic, count, average call duration, and corresponding department.  At this point, the data was ready for analysis.

\subsection{Duration}
Call duration by topics was calculated in the dataframe creation above, using the ``.groupby()'' method in Pandas.  However, when the dataframe was sorted to see the quickest and longest average duration, several topics with a low volume rose to the top, some with only one total call.  In a sense, such records can be viewed as outliers, and were thus removed from reporting in the final dashboard.  In the end, only topics with a total volume above the first quartile (of total calls) were included in this analysis.  By reducing the topics analyzed, the call center will be able to focus on topics that take a notable duration to resolve, but that also occur frequently enough to warrant attention.  This code, however, can be easily adjusted to a different threshold, if management so desires.

\subsection{Volume}
The total number of calls by topic was then analyzed.  The dataframe that was used in the previous analysis was again used for this.  Using the ``.groupby()" method in Pandas in conjunction with ``.count()", the total number of calls for each topic was appended to the dataframe.  The most frequent call topics were quickly identified, and many showed a significant overlap with the highest volume departments.  As a result, the volume by topic in and of itself would not serve much use to the city.  Therefore, the top 7 highest volume departments (identified in Figure 2) were isolated from the rest of the data.  The individual topics within these departments were then identified.  By approaching the analysis in this manner, the city can easily see the most popular topics in the busiest departments.  In order to visualize this data in the most effective manner, a jitterplot was incorporated into the final dashboard.  This plot can be seen in Figure 4.

\begin{figure}[h]
  \includegraphics[scale=0.3]{jitterplot.png}
  \caption{Jitterplot of call volume by topic for the busiest departments.  Hovertool is utilized to show topic.  The topic in diamond is not plotted to scale}
\end{figure}

@@

\begin{thebibliography}{9}

\bibitem{python}
Python Software Foundation. (2016). Python 3.6. https://docs.python.org/3.6/.

\bibitem{anaconda}
Anaconda. (2018). Anaconda Distribution 5.2. https://docs.anaconda.com/anaconda/.

\end{thebibliography}
%\includepdf{311CallCenterDashboard}
